Index measures for oak decline severity using phenotypic descriptors

### 2019-12-13 manuscript/manuscript.Rmd
composite descriptor descriptions complete

Machine learning is finding increasing utility in biological and geographical applications due to it's ability to learn complex patterns withing and between 
A popular algorithm is that of random forest due to it's versatility, relatively trivial training requirements and resilience against over fitting [@breiman2001] (refs). 
It is an ensemble machine learning method where forests of bootstrap sampled decision trees are grown, yielding both out of bag sample prediction accuracies and variable importance.
It is also able to simultaneously handle both continuous and categorical variables.
Unsupervised learning is also possible [@shi2006]

@@

Descriptors were collected relating to the trees shape, size, crown condition and the presence of biotic agents including insects, bacteria and fungi.
These descriptors are shown in Table #.
Crown radius, canopy closure and tap tests for dead stem tissue were measured for each cardinal point.
The crown radius was
The AOD stem symptoms (active bleeds, black staining, callused wounds) were collected by measuring longest axis of each individual bleed, stain or callused wound found within the basal 3m of the stem (ref).
The frequency of *A. biguttatus* exit holes were counted along with the presence or absence of each of oval shaped exit holes, small circular shaped exit holes and ground level fungal fruiting bodies were recorded from the basal 2m of the stem (ref). 
All trees were manually assigned a decline status category; control, AOD, COD or remission.

All raw data and analysis code used to generate this manuscript are available at https://github.com/jasenfinch/Index_measures_for_oak_decline_severity_using_phenotypic_descriptors.

@@

**Composite crown volume (m^3^)** - Estimated volume of crown as an paraboloid proportional to the crown condition. Based on the composite crown volumne from @zarnoch2004 but uses crown condition instead of crown density alone.

@@

**Composite crown surface area (m^2^)** -

@@

**Crown production efficiency** -

>>>>

Machine learning is finding increasing utility in biological and geographical applications due to it's ability to learn complex patterns within and between variables.
A popular algorithm is that of random forest due to it's versatility, relatively trivial training requirements and resilience against over fitting [@breiman2001] (refs). 
It is an ensemble machine learning method where forests of bootstrap sampled decision trees are grown, yielding both out of bag sample prediction accuracies and variable importance.
It is also able to simultaneously handle both continuous and categorical variables.
Unsupervised learning is also possible [@shi2006]

@@

Descriptors were collected relating to the trees shape, size, crown condition and the presence of biotic agents including insects, bacteria and fungi.
Crown radius, canopy closure and tap tests for dead stem tissue were measured for each cardinal point.
The crown radius was
The AOD stem symptoms (active bleeds, black staining, callused wounds) were collected by measuring longest axis of each individual bleed, stain or callused wound found within the basal 3m of the stem (ref).
The frequency of *A. biguttatus* exit holes were counted along with the presence or absence of each of oval shaped exit holes, small circular shaped exit holes and ground level fungal fruiting bodies were recorded from the basal 2m of the stem (ref). 
These descriptors are shown in Supplementary table #.
All trees were manually assigned a decline status category; control, AOD, COD or remission.
All raw data and analysis code used to generate this manuscript are available at https://github.com/jasenfinch/Index_measures_for_oak_decline_severity_using_phenotypic_descriptors.

@@

**Composite crown volume (m^3^)** - Estimated volume of crown as an paraboloid proportional to the crown condition. Adapted from the composite crown volume from @zarnoch2004 using crown condition instead of crown density alone.

@@

**Composite crown surface area (m^2^)** - Estimated crown surface area proportional to crown condition. Adapted from @laroque1994.

@@

**Crown production efficiency** - Estimate of crown production effieceny as a ratio between the composite crown surface area and the composite crown volumne [@zarnoch2004].
